AttCRISPR : an spacetime interpretable model for sgRNA efficiency prediction

### 2020-10-11 Paper/main.tex
10-11 paper

\begin{figure}[!tpb]%figure2
    \centerline{\includegraphics[width=86mm]{CNNv2.png}}
    \caption{The architecture of spatial domain method in AttCRISPR. 
    The input of the method is encoded sgRNA sequence, a $21\times 4$ one-hot matrix. 
    Then refine it through a spatial attention module, which could tell us the importance of a specific matrix element (or just say, pixel).
    Finally, a simple convolutional neural network is applied.}\label{fig:02}
\end{figure}
As Figure~2\vphantom{\ref{fig:02}} show, method in spatial domain relies on the convolutional neural network. 
As previously mentioned, sgRNA sequence has been encoded into a $21\times 4$ one-hot matrix $X_{oh}$, and we regard $X_{oh}$ as a binary image. 
Then, convolution kernels with different size are used to extract potential spatial features just like what other do in computer vision. 
According to the foregoing, the Spatial Attention Module proposed by \citeauthor{woo2018cbam:} can be applied in our method in spatial domain, 
which was used to improve the performance of CNN in vision tasks.

\begin{figure}[!tpb]%figure3
    \centerline{\includegraphics[width=86mm]{spatialmodule.png}}
    \caption{Details of spatial attention module. A convolution layer is used to generate multi-channel map from $X_{oh}$.
    Then concatenated the output of both max-pooling and average-pooling method and forward it to the last convolution layer.
    A sigmoid is used to map the final result to a range of zero to one.
    }\label{fig:03}
\end{figure}

As Figure~3\vphantom{\ref{fig:03}} show, Given an one-hot map $X_{oh}$, spatial attention module generating spatial attention matrix 
$A_s\in\mathbb{R}^{l\times 4}$ with the same as $X_{oh}$ in shape. 
Each element of $A_s$ is constrained to a range of zero to one, implemented by a sigmoid function, which reflects the importance of the corresponding elements of $X_{oh}$.
The overall spatial attention process can be summarized as:
\begin{equation}
\left\{\begin{matrix}
X_{mc} = f^{3\times4}(X_{oh})
\\A_s = \sigma(f^{3\times2}([AvgPool(X_{mc});MaxPool(X_{mc})]))
\\X_{rf} = A_s\otimes X_{oh}
\end{matrix}\right.\label{eq:11}
\end{equation}
where $f^{p\times q}$ represents a convolution operation with the filter size of $p\times q$, 
$p,q\in\mathbb{Z}^{+}$, $X_{mc}$ is a multi-channel map generated by $X_{oh}$, $\sigma$ denotes the sigmoid function, 
AvgPool denotes the average-pooling operation, MaxPool denotes the max-pooling operation, $\otimes$ denotes element-wise multiplication.
\begin{figure}[!tpb]%figure4
    \centerline{\includegraphics[width=86mm]{temporalmodule.png}}
    \caption{The architecture of temporal domain method in AttCRISPR. 
    The input of the method is embedding sgRNA sequence, a $21\times e$ embedding matrix, where $e$ is the dimension of a nucleotide embedding vector. 
    }\label{fig:04}
\end{figure}
Spatial attention matrix $A_s$ formally conforms to our proposed definition of first-order preference (each element of $X_{oh}$ is multiplied by the corresponding element of $A_s$), 
in other word, element of $A_s$ reveal how important the corresponding elements in $X_{oh}$ is. 
We think it can reveal the preference of the scoring function at each position. 
For instance, follow the encoding rules above, train spatial domain part of AttCRISPR with the WT-SpCas9 dataset. 
Then take the average of all spatial attention matrix, and the element in the first row and third column are closer to 1, which means when calculate the final score, 
G typically may have a important contribution at first position within sgRNA sequence. 
In fact, this corresponds to some early studies concerning the Human (hU6) promoter, which is believed to require G as the first nucleotide of its transcript \citep{cong2013multiplex,jinek2012a,mali2013rnaguided}.
\subsubsection{Method and attention in temporal domain}

\begin{figure}[!tpb]%figure3
    \centerline{\includegraphics[width=86mm]{temporalattentionmodule.png}}
    \caption{wait2write
    }\label{fig:03}

@@

@@

\subsubsection{Model ensemble following stacking strategy}

@@

The dataset we used for training, validation and testing is built by \citeauthor{wang2019optimized} \citep{wang2019optimized}. 
We extracted 55604, 58617, 56888 sgRNAs with activity (represented by insertion/deletion (indel)) for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1 respectively, from its source data. 
As far as we know, the best performance on this dataset is the DeepHF of \citeauthor{wang2019optimized} currently, which base on deep learning and bidirectional LSTM architecture. 
DeepHF achieved Spearman correlation coefficients of 0.867, 0.862 and 0.860 for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1 respectively. 
Another tool worth noting is CRISPRpred(SEQ), which is based on conventional machine learning and need complicated human feature engineering exercise \cite{MuhammadRafid2020}. 
They also test their tool with this dataset and achieved Spearman correlation coefficients of 0.838, 0.830 and 0.821 for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1 respectively (without any hyperparameter tuning).

@@

@@

Text
\begin{table}[!t]
    \processtable{This is table caption\label{Tab:01}} {\begin{tabular}{@{}llll@{}}\toprule head1 &
    head2 & head3 & head4\\\midrule
    row1 & row1 & row1 & row1\\
    row2 & row2 & row2 & row2\\
    row3 & row3 & row3 & row3\\
    row4 & row4 & row4 & row4\\\botrule
    \end{tabular}}{This is a footnote}
\end{table}

>>>>

\begin{figure}[!tpb]%figure2
    \centerline{\includegraphics[width=86mm]{CNNv2.png}}
    \caption{The architecture of spatial domain method in AttCRISPR. 
    The input of the method is encoded sgRNA sequence $X_{oh}$, a $21\times 4$ one-hot matrix. 
    Then refine it through a spatial attention module, which could tell us the importance of a specific matrix element (or just say, pixel).
    A simple convolutional neural network followed is applied to extract potential feature representation of sgRNA sequence.
    In the last step, we flatten the output of convolutional neural network into a one dimensional vector and 
    use a multilayer perceptron with sigmoid activation function to achieve the spatial output $y_s$.}\label{fig:02}
\end{figure}
As Figure~2\vphantom{\ref{fig:02}} show, method in spatial domain relies on the convolutional neural network. 
As previously mentioned, sgRNA sequence has been encoded into a $21\times 4$ one-hot matrix $X_{oh}$, and we regard $X_{oh}$ as a binary image. 
Then, convolution kernels with different size are used to extract potential spatial features just like what other do in computer vision. 
According to the foregoing, the Spatial Attention Module proposed by \citeauthor{woo2018cbam:} can be applied in our method in spatial domain, 
which was used to improve the performance of CNN in vision tasks.

\begin{figure}[!tpb]%figure3
    \centerline{\includegraphics[width=86mm]{spatialmodule.png}}
    \caption{Details of spatial attention module. A convolution layer is used to generate multi-channel map from $X_{oh}$.
    Then concatenated the output of both max-pooling and average-pooling method and forward it to the last convolution layer.
    A sigmoid function is used to map the final result to a range of zero to one at last, 
    which generates the spatial attention matrix $A_s$.
    }\label{fig:03}
\end{figure}

As Figure~3\vphantom{\ref{fig:03}} show, Given an one-hot map $X_{oh}$, spatial attention module generating spatial attention matrix 
$A_s\in\mathbb{R}^{l\times 4}$ with the same as $X_{oh}$ in shape. 
Each element of $A_s$ is constrained to a range of zero to one, implemented by a sigmoid function, which reflects the importance of the corresponding elements of $X_{oh}$.
The overall spatial attention process can be summarized as:
X_{mc} = f^{3\times4}(X_{oh})
\\A_s = \sigma(f^{3\times2}([AvgPool(X_{mc});MaxPool(X_{mc})]))
\\X_{rf} = A_s\otimes X_{oh}
\end{matrix}\right.\label{eq:11}
where $f^{p\times q}$ represents a convolution operation with the filter size of $p\times q$, 
$p,q\in\mathbb{Z}^{+}$, $X_{mc}$ is a multi-channel map generated by $X_{oh}$, $\sigma$ denotes the sigmoid function, 
AvgPool denotes the average-pooling operation, MaxPool denotes the max-pooling operation, $\otimes$ denotes element-wise multiplication.
\begin{figure}[!tpb]%figure4
    \centerline{\includegraphics[width=86mm]{temporalmodule.png}}
    \caption{The architecture of temporal domain method in AttCRISPR. 
    The input of the method is embedding sgRNA sequence $X_e$, a $21\times e$ embedding matrix, where $e$ is the dimension of a nucleotide embedding vector. 
    Then Keys $K$, Values $V$ and Queries $Q$ is generated through a classic encoder-decoder structure which is need for temporal attention module.
    }\label{fig:04}

@@

Spatial attention matrix $A_s$ formally conforms to our proposed definition of first-order preference (each element of $X_{oh}$ is multiplied by the corresponding element of $A_s$), 
in other word, element of $A_s$ reveal how important the corresponding elements in $X_{oh}$ is. 
We think it can reveal the preference of the scoring function at each position. 
For instance, follow the encoding rules above, train spatial domain part of AttCRISPR with the WT-SpCas9 dataset. 
Then take the average of all spatial attention matrix, and the element in the first row and third column are closer to 1, which means when calculate the final score, 
G typically may have a important contribution at first position within sgRNA sequence. 
In fact, this corresponds to some early studies concerning the Human (hU6) promoter, which is believed to require G as the first nucleotide of its transcript \citep{cong2013multiplex,jinek2012a,mali2013rnaguided}.
\subsubsection{Method and attention in temporal domain}

@@

\begin{figure}[!tpb]%figure3
    \centerline{\includegraphics[width=86mm]{temporalattentionmodule.png}}
    \caption{wait2write
    }\label{fig:03}
\end{figure}
\subsubsection{Model ensemble following stacking strategy}

@@

The dataset we used for training, validation and testing is built by \citeauthor{wang2019optimized} \citep{wang2019optimized}. 
We extracted 55604, 58617, 56888 sgRNAs with activity (represented by insertion/deletion (indel)) for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1 respectively, from its source data. 
As far as we know, the best performance on this dataset is the DeepHF of \citeauthor{wang2019optimized} currently, which base on deep learning and bidirectional LSTM architecture. 
DeepHF achieved Spearman correlation coefficients of 0.867, 0.862 and 0.860 for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1 respectively. 
Another tool worth noting is CRISPRpred(SEQ), which is based on conventional machine learning and need complicated human feature engineering exercise \citep{MuhammadRafid2020}. 
They also test their tool with this dataset and achieved Spearman correlation coefficients of 0.838, 0.830 and 0.821 for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1 respectively (without any hyperparameter tuning).
\citeauthor{wang2019optimized} implemente a tenfold shuffled validation to evaluate the stability of the performance of DeepHF.
To be more specific, each set is shuffled and divided into three parts, 76.5\%, 15\% and 8.5\% of the relevant data was used as the training, test and validation set respectively in a single experiment. 
The experiment is repeated ten times with the results recorded and averaged finally.
In order to make the comparison apples to apples, we will follow this strategy to design experiments.

@@

\begin{table}[!t]
    \processtable{This is table caption\label{Tab:01}} {\begin{tabular}{@{}llll@{}}\toprule head1 &
    head2 & head3 & head4\\\midrule
    row1 & row1 & row1 & row1\\
    row2 & row2 & row2 & row2\\
    row3 & row3 & row3 & row3\\
    row4 & row4 & row4 & row4\\\botrule
    \end{tabular}}{This is a footnote}
\end{table}

@@
